---
title: "Band Gaze"
output: html_notebook
---

```{r}
library(tidyverse)
library(ggthemes)
library(scales)
library(forcats)
```

```{r}
library(jpeg)
library(knitr)
```


```{r}
filename <- 'data/faces.csv'
faces <- read_csv(filename)

```

```{r}
faces <- faces %>% mutate(has_face = face_count > 0)
with_faces <- faces %>% filter(face_count > 0)
```

```{r}
unique_artists <- faces %>% distinct(cat_id, artist_id, .keep_all = TRUE)
```


## Artist Counts


```{r}
unique_artists %>%
  ggplot(aes(x = fct_rev(fct_infreq(cat_name)))) +
  coord_flip() +
  geom_bar() +
  scale_y_continuous(labels = comma) +
  labs(title = "Artists Per Category", y = "") +
  theme_fivethirtyeight()
```

```{r}
byCat <- unique_artists %>% group_by(cat_name) %>% mutate(cat_count = n()) %>%
  arrange(-cat_count) %>%
  group_by(cat_name, has_face) %>% summarise(count = n(), per = count / cat_count[1], cat_count = cat_count[1]) %>% arrange(cat_count)
```

```{r}
byCat %>% filter(has_face == TRUE) %>%
  ggplot(aes(x = fct_inorder(cat_name), y = count )) + 
  coord_flip() +
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = comma) +
  labs(title = "Count of Artists with Faces", y = "") +
  theme_fivethirtyeight()

```

```{r}
byCat %>% 
  ggplot(aes(x = fct_inorder(cat_name), y = per, fill = has_face)) + 
  coord_flip() +
  scale_y_continuous(labels = percent) +
  geom_bar(stat = "identity") + 
  labs(x = "", y = "", title = "Percent Artists with Face by Category") +
  theme_fivethirtyeight()

```


## Face Counts

```{r}
faces %>%
  ggplot(aes(x = (face_count))) +
  geom_histogram(binwidth = 1.0) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(0, 15, 1)) +
  theme_fivethirtyeight() +
  labs(title = "Distribution of Face Count", x = "Number of Faces in Artist Image")

```

```{r}
unique_artists %>%
  ggplot(aes(x = (face_count))) +
  geom_histogram(binwidth = 1.0) +
  scale_y_continuous(labels = comma) +
  scale_x_continuous(breaks = seq(0, 15, 1)) +
  theme_fivethirtyeight() +
  labs(title = "Distribution of Face Count for Artists", x = "Number of Faces in Artist Image")

```

Lots of no-faces detected and solo artists.

```{r}
max(unique_artists$face_count)
```



### Face Count by Genre

```{r}
byCatFaceCount <- unique_artists %>% filter(has_face == TRUE) %>%
  group_by(cat_name) %>% summarise(mean_face_count = mean(face_count), median_face_count = median(face_count))
  
```

```{r}

```


```{r}
unique_artists %>% filter(has_face == TRUE) %>% #filter(cat_id == "metal") %>%
  ggplot(aes(x = face_count)) +
  #geom_histogram(binwidth = 1.0) +
  #geom_histogram(aes(y=..count../sum(..count..)), binwidth = 1.0) + 
  geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]), binwidth = 1.0) + 
  scale_y_continuous(labels = percent) +
  scale_x_continuous(breaks = seq(1, 14, 2)) +
  facet_wrap(~ cat_name) + 
  labs(title = "Distribution of Face Count for Artists by Category", x = "Number of Faces in Artist Image") +
  theme_fivethirtyeight()
```

Cool!

So Country and Hip Hop artists almost always have solo images.

But you can't have a metal, punk, or rock band without at least a few people. 


## Gender

```{r}
with_faces %>%
  ggplot(aes(x = 1, fill = face_gender)) + 
  scale_y_continuous(labels = comma) +
  scale_x_continuous(labels = c()) +
  geom_bar() + 
  labs(x = "", y = "", title = "Count of Faces by Gender") +
  theme_fivethirtyeight()
```


```{r}
with_faces %>%
  ggplot(aes(x = 1, fill = face_gender)) + 
  scale_y_continuous(labels = percent) +
  scale_x_continuous(labels = c()) +
  geom_bar(aes(y = (..count..)/sum(..count..)), position = "dodge") + 
  labs(x = "", y = "", title = "Percent of Faces by Gender") +
  theme_fivethirtyeight()
```

Lots of Dudes in this data!

### Gender by Category


```{r}
with_faces %>% #filter(face_gender == 'female') %>% 
  ggplot(aes(x = face_gender))  +
  geom_bar() +
  #geom_histogram(aes(y=..count../sum(..count..)), binwidth = 1.0) + 
  #geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]), binwidth = 1.0) + 
  scale_y_continuous(labels = comma) +
  #scale_x_continuous(breaks = seq(1, 14, 2)) +
  facet_wrap(~ cat_name) + 
  labs(title = "Gender Counts of Faces by Category", x = "") +
  theme_fivethirtyeight()
```

```{r}
with_faces %>% #filter(face_gender == 'female') %>% 
  ggplot(aes(x = face_gender))  +
  #geom_bar() +
  #geom_histogram(aes(y=..count../sum(..count..)), binwidth = 1.0) + 
  geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]) ) + 
  scale_y_continuous(labels = percent) +
  #scale_x_continuous(breaks = seq(1, 14, 2)) +
  facet_wrap(~ cat_name) + 
  labs(title = "Gender Percents of Faces by Category", x = "") +
  theme_fivethirtyeight()
```


## Glasses

```{r}
with_faces %>%
  ggplot(aes(x = face_glasses, fill = face_glasses)) + 
  scale_y_continuous(labels = percent) +
  geom_bar(aes(y = (..count..)/sum(..count..)), position = "dodge") + 
  labs(x = "", y = "", title = "Percentage of Glasses on Faces", fill = "Glasses Type") +
  theme_fivethirtyeight()
```

Swimming Goggles ?? That can't be right.

```{r}
face_ind <- c(1,2,3)
filtered_faces <- with_faces %>% filter(face_glasses == 'SwimmingGoggles') %>% filter(face_count == 1) 
row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```


Well, ok. Maybe not.

```{r}
with_faces %>% filter(!(face_glasses == 'NoGlasses')) %>% filter(!(face_glasses == 'SwimmingGoggles')) %>%
  ggplot(aes(x = face_glasses, fill = face_glasses)) + 
  scale_y_continuous(labels = percent) +
  geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]) ) + 
  facet_wrap(~ cat_name) + 
  labs(x = "", y = "", title = "Sun Glasses vs Reading Glasses by Category", fill = "Glasses Type") +
  theme_fivethirtyeight()
```

## 3D Face Pose

There are 3 values recorded for face pose: Yaw, Roll, and Pitch. 

Most explainations of these features only explain it for [Aircraft](https://www.youtube.com/watch?v=pQ24NtnaLl8), but here is my understanding for human faces:

* Yaw: The direction your nose is pointing, to the left or right.
* Roll: Cocking your head to the left or right.
* Pitch: Looking up or down.

Unfortunately, Pitch isn't yet computed - so we can only look 

Lets look at Yaw and Roll:

### Yaw

```{r}
with_faces %>%
  ggplot(aes(x = face_yaw)) +
  geom_histogram(binwidth = 2) +
  scale_y_continuous(labels = comma) +
  #scale_x_continuous(breaks = seq(0, 15, 1)) +
  theme_fivethirtyeight() +
  labs(title = "Distribution of Yaw Metric", x = "")
```

Mean:

```{r}
mean(with_faces$face_yaw)
```

Median:

```{r}
median(with_faces$face_yaw)
```

So there is a slight trend of the nose pointing to the left ... or at least that is what the algorithm is detecting. Could be a bias!

But really, as we see below, -1 or -2 is really not visibly noticable as a head turn.


### Yaw Examples

**Light Negative Yaw**

Faces with yaw between -1 and -1.8

```{r}
face_ind <- c(1,2,3)
filtered_faces <- with_faces %>% filter(face_count == 1) %>% filter(face_yaw < -1) %>% filter(face_yaw > -1.8)
row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```

**Medium Negative Yaw**

Faces with Yaw between -10 and -15

```{r}
med_neg_yaw <- with_faces %>% filter(face_count == 1) %>% filter(face_yaw < -10) %>% filter(face_yaw > -15)
row <- med_neg_yaw[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```

**Medium Positive Yaw**

Faces with Yaw between 10 and 15

```{r}
med_pos_yaw <- with_faces %>% filter(face_count == 1) %>% filter(face_yaw > 10) %>% filter(face_yaw < 15)
row <- med_pos_yaw[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```

**Heavy Positive Yaw**

Faces with > 30 Yaw

```{r}
filtered_faces <- with_faces %>% filter(face_count == 1) %>% filter(face_yaw > 30)
row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```


### Roll

Let's get our Roll on.

```{r}
with_faces %>%
  ggplot(aes(x = face_roll)) +
  geom_histogram(binwidth = 2) +
  scale_y_continuous(labels = comma) +
  #scale_x_continuous(breaks = seq(0, 15, 1)) +
  theme_fivethirtyeight() +
  labs(title = "Distribution of Roll Metric", x = "")
```


```{r}
mean(with_faces$face_roll)
```

```{r}
median(with_faces$face_roll)
```

Roll looks to be distributed pretty evenly - but there is a longer tail to the right. The summary stats indicate there is a slight drag to the left.   

Lets look at some examples!

### Roll Examples

**Light Negative Roll**

Faces with roll between -1 and -2

```{r}
face_ind <- c(1,2,3)
filtered_faces <- with_faces %>% filter(face_count == 1) %>% filter(face_roll < -1) %>% filter(face_roll > -2)
row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```

**Heavy Negative Roll**

Faces with roll < -20

```{r}
filtered_faces <- with_faces %>% filter(face_count == 1) %>% filter(face_roll < -20)
row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```

**Heavy Positive Roll**

Faces with roll > 20

```{r}
filtered_faces <- with_faces %>% filter(face_count == 1) %>% filter(face_roll > 20)
row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```


## Facial Hair

Look just at Beards for a second:

```{r}
with_faces %>% filter(face_gender == 'male') %>%
  ggplot(aes(x = face_beard)) + 
  geom_freqpoly(binwidth = 0.01) + 
  labs(title = "Male Beard Likelihood Values")
```

Interesting! So this isn't really a continuous variable, instead its descritized. 

To double check that, here are the unique values of `face_beard`:

```{r}
sort(unique(with_faces$face_beard))
```



```{r}
facial_hair <- with_faces %>% tidyr::gather(key = hair_type, value = hair_ratio, face_beard, face_moustache, face_sideburns)
```


```{r}
facial_hair %>% filter(face_gender == 'male') %>%
  ggplot(aes(x = as.factor(hair_ratio), fill = hair_type)) + 
  geom_bar(position = "dodge") + 
  scale_y_continuous(labels = comma) +
  labs(title = "Male Facial Hair Counts", fill = "Hair Type") + 
  theme_fivethirtyeight()
```

```{r}
facial_hair %>% filter(face_gender == 'male') %>%
  ggplot(aes(x = as.factor(hair_ratio), fill = hair_type)) + 
  #geom_bar(position = "dodge") + 
  geom_bar(aes(y=..count../sum(..count..)), position = "dodge") + 
  scale_y_continuous(labels = percent) +
  labs(title = "Male Facial Hair Percents", fill = "Hair Type") + 
  theme_fivethirtyeight()
```

```{r}

facial_hair %>% filter(face_gender == 'male') %>%
  ggplot(aes(x = as.factor(hair_ratio), fill = hair_type)) + 
  #geom_bar(position = "dodge") + 
  geom_bar(aes(y=..count../sum(..count..)), position = "dodge") + 
  scale_y_continuous(labels = percent) +
  labs(title = "Male Facial Hair Percents", fill = "Hair Type") + 
  facet_wrap(~ hair_type) +
  theme_fivethirtyeight()
```


What about `female` identified faces?

```{r}
facial_hair %>% filter(face_gender == 'female') %>%
  ggplot(aes(x = hair_ratio, fill = hair_type)) + 
  geom_histogram(binwidth = 0.01) +
  labs(title = "Women don't have Facial Hair in this Dataset / Algorithm")
```

### Facial Hair by Category





```{r}

facial_hair %>% filter(face_gender == 'male') %>% filter(hair_type == 'face_beard') %>%
  ggplot(aes(x = as.factor(hair_ratio) )) + 
  #geom_bar(position = "dodge") + 
  #geom_bar(aes(y=..count../sum(..count..))) + 
  geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
  scale_y_continuous(labels = percent) +
  labs(title = "Beards by Category") + 
  facet_wrap(~ cat_name) +
  theme_fivethirtyeight()
```

```{r}
filtered_faces <- with_faces %>% filter(face_count == 1) %>% filter(face_gender == 'male') %>% filter(face_beard == 1.0) 

row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```


```{r}
filtered_faces <- with_faces %>% filter(face_count == 1) %>% filter(face_gender == 'male') %>% filter(face_beard == 1.0) 

row <- filtered_faces[face_ind,]
filename <- paste("data/imgs/", row$cat_id, "/", row$artist_id, ".jpg", sep="")
knitr::include_graphics(filename)
```

```{r}
male_hair <- with_faces %>% filter(face_gender == 'male') %>% mutate(has_facial_hair = face_beard + face_moustache > 0.4)
```


```{r}
male_hair %>%
  ggplot(aes(x = has_facial_hair, fill = has_facial_hair))  +
  #geom_bar() +
  #geom_histogram(aes(y=..count../sum(..count..)), binwidth = 1.0) + 
  geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]) ) + 
  scale_y_continuous(labels = percent) +
  #scale_x_continuous(breaks = seq(1, 14, 2)) +
  facet_wrap(~ cat_name) + 
  labs(title = "Gender Percents of Faces by Category", x = "") +
  theme_fivethirtyeight()
```

## Smiles

Most artists appear not to be smiling.

```{r}
with_faces %>% 
  ggplot(aes(x = face_smile)) + 
  geom_histogram(binwidth = 0.1) +
  labs(title = "Smile Distribution")
```

Gender doesn't seem to matter much:

```{r}
with_faces %>% 
  ggplot(aes(x = face_smile, fill = face_gender)) + 
  geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]),binwidth = 0.1 ) + 
  #geom_histogram(aes(y=..count../sum(..count..)), binwidth = 0.1, position = "dodge") +
  facet_wrap(~ face_gender) +
  scale_y_continuous(labels = percent) +
  labs(title = "Smile Distribution by Gender", y = "", x = "")
```





```{r}
with_faces %>% mutate(is_smiling = face_smile > 0.25) %>%
  ggplot(aes(x = is_smiling, fill = is_smiling)) + 
  geom_bar(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..]) ) + 
  #geom_histogram(aes(y=..count../sum(..count..)), binwidth = 0.1, position = "dodge") +
  facet_wrap(~ face_gender) +
  scale_y_continuous(labels = percent) +
  labs(title = "Smile Distribution by Gender", y = "", x = "")
```

## Image Sizes


```{r}
img_details_filename <- "data/images.csv"
images <- read_csv(img_details_filename)
```

```{r}
images %>%
  ggplot(aes(x = img_width)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Image Widths")
  
```

```{r}
summary(images$img_width)
```


```{r}
images %>%
  ggplot(aes(x = img_height)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Image Height")
  
```


```{r}
summary(images$img_height)
```

```{r}
images %>%
  ggplot(aes(x = img_width, y = img_height)) +
  geom_point( alpha = 1 / 10) +
  stat_density2d(aes(fill = ..level..), geom="polygon")
```

## Face Image Sizes

```{r}
face_details_filename <- "data/face_sizes.csv"
face_details <- read_csv(face_details_filename)
```





```{r}
face_details %>%
  ggplot(aes(x = img_width, y = img_height)) +
  geom_point( alpha = 1 / 10) +
  stat_density2d(aes(fill = ..level..), geom="polygon")
```


```{r}
face_details_by_artist <- face_details %>% group_by(cat_id, artist_id) %>% mutate(face_count = n()) %>% ungroup()
```


```{r}
face_details_by_artist %>% distinct(cat_id, artist_id, .keep_all = TRUE) %>%
  ggplot(aes(x = as.factor(face_count))) + 
  geom_bar() + 
  labs(title = "Counts of Artists by Face Count")
```


```{r}
face_details_by_artist %>% 
  ggplot(aes(x = as.factor(face_count), y = img_width)) + 
  geom_boxplot() + 
  labs(title = "Face Width by Face Count")
```

```{r}
face_details_by_artist %>% 
  ggplot(aes(x = as.factor(face_count), y = img_height)) + 
  geom_boxplot() + 
  labs(title = "Face Height by Face Count")

```


```{r}

face_details_by_artist %>% 
  ggplot(aes(x = as.factor(face_count), y = img_height)) + 
  geom_boxplot() + 
  facet_wrap(~ cat_id) +
  labs(title = "Face Height by Face Count")
```




